Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-09-01 BD_radvel/astro-ph/kurosawa.tex
I forgot to add some files last time.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@813 25464f21-6099-402c-b30a-640f1d40d38b

The data were reduced via the standard ESO pipeline procedures for
UVES echelle spectra. In summary, the data were corrected for bias,
interorder background, sky background, sky emission lines and cosmic
ray hits. They were then flattened, optimally extracted, and finally
the different orders were merged. No binning was performed to achieve
high resolution required for the RV measurements. The wavelength was
calibrated using the Thorium-Argon arc spectra with a typical value
of the standard deviation of the dispersion solution of 5~m\AA\,
which corresponds to 0.2~$\kmps$ at the central wavelength 8600~\AA.
However, the autoguiding of the telescope keeps the star at the centre
of the slit with about a tenth of the FWHM ($1\,\kmps$) which sets
the upper limit for the systematic error in the RV measurements
\citep{bailer-jones:2004}.
\textcolor{red}{
In fact, we find that any systematic error is negligible compared to
the random error associated with our radial velocity measurements (see
Section~\ref{sub:Radial-velocities}).
}
A typical
signal-to-noise ratio (S/N) per wavelength bin of the spectra is about
15, and the heliocentric velocity correction was applied to the final
spectra.

\textcolor{red}{
In addition to the main targets in Table~\ref{tab:literatures}, we
also obtained the spectra of LHS~49 (on 2004-April-17) and HD~140538
(on 2004-May-06), which are the RV template and the RV standard
stars respectively.  These data were obtained using the same telescope
and the same instrument setup as for our main targets.  The S/N of
the both objects were about 60. 
}

@@

\caption{Summary of the observations, the heliocentric radial velocities ($\mathrm{RV}$)
from two-epoch and the average rotational velocities ($v\sin i$).
The uncertainties of relative radial velocities ($\sigma_{\mathrm{RRV}}$)
with respect to the template star LHS~049 and the average radial
velocities ($\overline{\mathrm{RV}}$) are also given.  
\textcolor{red}{
N.B.~the uncertainties listed along with the heliocentric radial
velocities in column~4 and 6 include the uncertainty in the
heliocentric RV of the template star, but $\sigma_{\mathrm{RRV}}$ in
column~5 do not.
}
}

@@

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard
HD~140538 \textcolor{red}{(G2.5V)}
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed \'{e}celle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric 
\textcolor{red}{
$\mathrm{RV_{HD\,140538}=}19.17\pm0.40\,\kmps$
}
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep*{udry:1999}.
\textcolor{red}{
  Note that the uncertainty in $\mathrm{RV_{HD\,140538}}$ was computed
  by combining (in quadrature) the uncertainty ($0.4~\kmps$) in the
  heliocentric RV of the template star HD~74497 (G3V) and the
  uncertainty in the relative RV of HD~140538 ($0.02~\kmps$), measured
  by cross correlation with the template star.  Clearly, the former
  dominates in the uncertainty of
  $\mathrm{RV_{HD\,140538}}$ shown above. The uncertainty in the
  heliocentric RV of the template is rather large because it was
  determined from a few photospheric lines.
}

\textcolor{red}{
  The results of the heliocentric RV measurements (from two epochs for
  each object) are summarised in Table 2 along with their
  uncertainties.  The uncertainties listed in columns~4 and 6 include
  the uncertainty of the absolute RV of the template star, added in
  quadrature with the uncertainty found from the cross correlation
  analysis. On the other hand, the uncertainties in the relative
  radial velocities ($\sigma_{\mathrm{RRV}}$) listed in column~5 of
  the table \emph{do not} include the uncertainty of the template star. It is
  $\sigma_{\mathrm{RRV}}$ that is relevant for the identification of
  RV variable objects.
}

To determine the $\sigma_{\mathrm{RRV}}$, we have used the following
procedure: (1)~add random gaussian deviate noise to the object spectra using the
corresponding variance spectra\footnote{\textcolor{red}{Defined by the variance of the
$\chi^{2}$ fit to the signal obtained during optimum extraction \citep[c.f.][]{horne:1986}.}}, 
(2)~compute the cross-correlation
curve using the spectra with added noise and the template, (3)~measure
the RV by locating the peak, (4)~repeat 1--3 for 100 times, and compute the standard
deviation of the 100 RV values. This should provide a good estimate of
uncertainties in RVs cased by the uncertainties in the flux levels.
We also estimated the RV uncertainties using the standard deviations on the
mean of the two independent RV measurements obtained from the two
consecutive spectra taken on the same night (c.f.~\citealt{joergens:2006}).
We found that these estimates agreed well with those obtained from
the Monte Carlo method. The `average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in Table~\ref{tab:summary_results}.

For each object and for each RV measurement, the deviations ($\Delta\mathrm{RV}$)
from the average RV are computed and summarized in Fig.~\ref{fig:radvel01}
along with their uncertainties ($\sigma_{\mathrm{RRV}}$) in order
to aid the identification of multiplicity candidates. Note that in
computing $\Delta\mathrm{RV}$ we do not require the knowledge of
absolute\textcolor{red}{/heliocentric} radial velocities, but only the relative
velocities (with respect to a template). 
We also  cross-correlated object spectra from the two epochs 
(instead of using the spectra of the template LHS~49)  in order to find
$\Delta\mathrm{RV}$. The resulting cross-correlation functions were
found to be too noisy (due to the relatively low S/N in the object
spectra) for the  peak positions of the cross-correlation curves to be
reliably located.

To identify an object with a RV variation with a statistical
significance from our sample, we apply the method described by
\citet{maxted:2005}, which we briefly summarize next, to our data.
There are three steps in this method: 
(1)~compute the $\chi^{2}$ by fitting the two-epoch RV data for each 
object with a constant function (a zero-th order polynomial),
(2)~compute the corresponding $\chi^{2}$ probability ($p$),
(3)~designate the object as a non-constant RV object or a binary
candidate if $p<10^{-3}$ (\textcolor{red}{0.1~per~cent}).  When computing $\chi^{2}$, we
use the uncertainties in the relative RV ($\sigma_{\mathrm{RRV}}$). 
The number of degree of freedom in the fitting procedure is obviously
1.  A similar method was also used in a recent RV survey of VLM stars
by \citet{basri:2006}.

We have computed $p$ for all the objects (Table~\ref{tab:chisq_prob}),
and have plotted the results  
as a histogram of $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on the right (with the large $-\log{p}$ values)
occupied by the RV variable candidates.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. 
\textcolor{red}{
To quantify this point, we compute the
$\chi^{2}$ probability to test the goodness of the fit of the expected
distribution to the histogram. We restrict the fit to the first 4 bins
from the left $\left(0 < -\log_{10}p < 2 \right)$ of the
histogram since the objects in the second bin from the right
$\left(2.5 < -\log_{10}p < 3 \right)$ are `potentially'
RV-variable, although we have flagged them as the RV-constant
objects. In this analysis, we find the probability of 0.5 which
indicates that the fit is reasonable. Further, if we re-normalized the
expected distribution, accounting for that fact we neglected the
objects in the second bin from the right, we find the $\chi^{2}$
probability of the fit increases to 0.9.  This, in turn, indicates
that our uncertainties in RV values are reasonable, and an additional
systematic error is not necessary in this analysis. 
}
\textcolor{red}{
This assertion is supported by the re-analysis of BD RVs, obtained by
\citet{joergens:2006} using the same instruments setup
as ours, by \citet{maxted:2005}. 
}

@@

\caption{Histogram of the $\chi^2$ probability ($p$) for fitting the
  observed (relative) RV values with a constant (a horizontal
  line). The constant used in the fit is determined from the weighted mean of
  the two RV measurements for each object.  The objects with
  $p<10^{-3}$ are identified as \emph{non-constant} or \emph{multiple},
  which appear in the right-most bin in the histogram. There
  are four objects which satisfy this condition. 
  \textcolor{red}{The expected distribution (dashed) for RV constant
  objects is also shown for a comparison, and is 
  normalized to the total number of objects that have $p>10^{-3}$}.
  The match between the expected distribution and the
  histogram is reasonable \textcolor{red}{(c.f.~Section~\ref{sub:Radial-velocities})},
  indicating that our uncertainty estimates 
  in RV values are also reasonable.  }

@@

Follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates identified in earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} points out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straightforwardly,
i.e. without correcting for bias. 
\textcolor{red}{
It is possible that our binary fraction is an overestimate since we
are biased to brighter objects. However the lack of a double-line
system suggests that a primary-to-secondary luminosity ratio of $\gg 1$
for the RV variables found in this paper. 
}

>>>>

The data were reduced via the standard ESO pipeline procedures for
UVES echelle spectra. In summary, the data were corrected for bias,
interorder background, sky background, sky emission lines and cosmic
ray hits. They were then flattened, optimally extracted, and finally
the different orders were merged. No binning was performed to achieve
high resolution required for the RV measurements. The wavelength was
calibrated using the Thorium-Argon arc spectra with a typical value
of the standard deviation of the dispersion solution of 5~m\AA\,
which corresponds to 0.2~$\kmps$ at the central wavelength 8600~\AA.
However, the autoguiding of the telescope keeps the star at the centre
of the slit with about a tenth of the FWHM ($1\,\kmps$) which sets
the upper limit for the systematic error in the RV measurements
\citep{bailer-jones:2004}.
\textcolor{black}{
In fact, we find that any systematic error is negligible compared to
the random error associated with our radial velocity measurements (see
Section~\ref{sub:Radial-velocities}).
}
A typical
signal-to-noise ratio (S/N) per wavelength bin of the spectra is about
15, and the heliocentric velocity correction was applied to the final
spectra.

\textcolor{black}{
In addition to the main targets in Table~\ref{tab:literatures}, we
also obtained the spectra of LHS~49 (on 2004-April-17) and HD~140538
(on 2004-May-06), which are the RV template and the RV standard
stars respectively.  These data were obtained using the same telescope
and the same instrument setup as for our main targets.  The S/N of
the both objects were about 60. 
}

@@

\caption{Summary of the observations, the heliocentric radial velocities ($\mathrm{RV}$)
from two-epoch and the average rotational velocities ($v\sin i$).
The uncertainties of relative radial velocities ($\sigma_{\mathrm{RRV}}$)
with respect to the template star LHS~049 and the average radial
velocities ($\overline{\mathrm{RV}}$) are also given.  
\textcolor{black}{
N.B.~the uncertainties listed along with the heliocentric radial
velocities in column~4 and 6 include the uncertainty in the
heliocentric RV of the template star, but $\sigma_{\mathrm{RRV}}$ in
column~5 do not.
}
}

@@

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard
HD~140538 \textcolor{black}{(G2.5V)}
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed \'{e}celle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric 
\textcolor{black}{
$\mathrm{RV_{HD\,140538}=}19.17\pm0.40\,\kmps$
}
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep*{udry:1999}.
\textcolor{black}{
  Note that the uncertainty in $\mathrm{RV_{HD\,140538}}$ was computed
  by combining (in quadrature) the uncertainty ($0.4~\kmps$) in the
  heliocentric RV of the template star HD~74497 (G3V) and the
  uncertainty in the relative RV of HD~140538 ($0.02~\kmps$), measured
  by cross correlation with the template star.  Clearly, the former
  dominates in the uncertainty of
  $\mathrm{RV_{HD\,140538}}$ shown above. The uncertainty in the
  heliocentric RV of the template is rather large because it was
  determined from a few photospheric lines.
}

\textcolor{black}{
  The results of the heliocentric RV measurements (from two epochs for
  each object) are summarised in Table 2 along with their
  uncertainties.  The uncertainties listed in columns~4 and 6 include
  the uncertainty of the absolute RV of the template star, added in
  quadrature with the uncertainty found from the cross correlation
  analysis. On the other hand, the uncertainties in the relative
  radial velocities ($\sigma_{\mathrm{RRV}}$) listed in column~5 of
  the table \emph{do not} include the uncertainty of the template star. It is
  $\sigma_{\mathrm{RRV}}$ that is relevant for the identification of
  RV variable objects.
}

To determine the $\sigma_{\mathrm{RRV}}$, we have used the following
procedure: (1)~add random gaussian deviate noise to the object spectra using the
corresponding variance spectra\footnote{\textcolor{black}{Defined by the variance of the
$\chi^{2}$ fit to the signal obtained during optimum extraction \citep[c.f.][]{horne:1986}.}}, 
(2)~compute the cross-correlation
curve using the spectra with added noise and the template, (3)~measure
the RV by locating the peak, (4)~repeat 1--3 for 100 times, and compute the standard
deviation of the 100 RV values. This should provide a good estimate of
uncertainties in RVs cased by the uncertainties in the flux levels.
We also estimated the RV uncertainties using the standard deviations on the
mean of the two independent RV measurements obtained from the two
consecutive spectra taken on the same night (c.f.~\citealt{joergens:2006}).
We found that these estimates agreed well with those obtained from
the Monte Carlo method. The `average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in Table~\ref{tab:summary_results}.

For each object and for each RV measurement, the deviations ($\Delta\mathrm{RV}$)
from the average RV are computed and summarized in Fig.~\ref{fig:radvel01}
along with their uncertainties ($\sigma_{\mathrm{RRV}}$) in order
to aid the identification of multiplicity candidates. Note that in
computing $\Delta\mathrm{RV}$ we do not require the knowledge of
absolute\textcolor{black}{/heliocentric} radial velocities, but only the relative
velocities (with respect to a template). 
We also  cross-correlated object spectra from the two epochs 
(instead of using the spectra of the template LHS~49)  in order to find
$\Delta\mathrm{RV}$. The resulting cross-correlation functions were
found to be too noisy (due to the relatively low S/N in the object
spectra) for the  peak positions of the cross-correlation curves to be
reliably located.

To identify an object with a RV variation with a statistical
significance from our sample, we apply the method described by
\citet{maxted:2005}, which we briefly summarize next, to our data.
There are three steps in this method: 
(1)~compute the $\chi^{2}$ by fitting the two-epoch RV data for each 
object with a constant function (a zero-th order polynomial),
(2)~compute the corresponding $\chi^{2}$ probability ($p$),
(3)~designate the object as a non-constant RV object or a binary
candidate if $p<10^{-3}$ (\textcolor{black}{0.1~per~cent}).  When computing $\chi^{2}$, we
use the uncertainties in the relative RV ($\sigma_{\mathrm{RRV}}$). 
The number of degree of freedom in the fitting procedure is obviously
1.  A similar method was also used in a recent RV survey of VLM stars
by \citet{basri:2006}.

We have computed $p$ for all the objects (Table~\ref{tab:chisq_prob}),
and have plotted the results  
as a histogram of $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on the right (with the large $-\log{p}$ values)
occupied by the RV variable candidates.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. 
\textcolor{black}{
To quantify this point, we compute the
$\chi^{2}$ probability to test the goodness of the fit of the expected
distribution to the histogram. We restrict the fit to the first 4 bins
from the left $\left(0 < -\log_{10}p < 2 \right)$ of the
histogram since the objects in the second bin from the right
$\left(2.5 < -\log_{10}p < 3 \right)$ are `potentially'
RV-variable, although we have flagged them as the RV-constant
objects. In this analysis, we find the probability of 0.5 which
indicates that the fit is reasonable. Further, if we re-normalized the
expected distribution, accounting for that fact we neglected the
objects in the second bin from the right, we find the $\chi^{2}$
probability of the fit increases to 0.9.  This, in turn, indicates
that our uncertainties in RV values are reasonable, and an additional
systematic error is not necessary in this analysis. 
}
\textcolor{black}{
This assertion is supported by the re-analysis of BD RVs, obtained by
\citet{joergens:2006} using the same instruments setup
as ours, by \citet{maxted:2005}. 
}

@@

\caption{Histogram of the $\chi^2$ probability ($p$) for fitting the
  observed (relative) RV values with a constant (a horizontal
  line). The constant used in the fit is determined from the weighted mean of
  the two RV measurements for each object.  The objects with
  $p<10^{-3}$ are identified as \emph{non-constant} or \emph{multiple},
  which appear in the right-most bin in the histogram. There
  are four objects which satisfy this condition. 
  \textcolor{black}{The expected distribution (dashed) for RV constant
  objects is also shown for a comparison, and is 
  normalized to the total number of objects that have $p>10^{-3}$}.
  The match between the expected distribution and the
  histogram is reasonable \textcolor{black}{(c.f.~Section~\ref{sub:Radial-velocities})},
  indicating that our uncertainty estimates 
  in RV values are also reasonable.  }

@@

Follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates identified in earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} points out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straightforwardly,
i.e. without correcting for bias. 
\textcolor{black}{
It is possible that our binary fraction is an overestimate since we
are biased to brighter objects. However the lack of a double-line
system suggests that a primary-to-secondary luminosity ratio of $\gg 1$
for the RV variables found in this paper. 
}
